\myTitle

### 2024-06-18 tex/body-text-of-paper.tex

\subsubsection*{Markers around the RoSA region of \citet{thompson2020complex}}

We include in the reference baseline amplicons targeting SNPs within the ``Region of Strongest Assocation'' (RoSA)
between spring-run and fall-run fish on Chromosome Ots28 described in \citet{thompson2020complex}.  These are useful markers to
have in the panel, as typing them routinely will create a data base for understanding the spatial and temporal
distribution of the alleles at this region; however we have omitted these markers for calculating likelihoods for
GSI, since their distribution across populations in the Feather River has been clearly and evidently
influenced by hatchery practices (see Results), and it is not necessary to include them for accurate assignment.

@@

We calculated \citet{weir1984estimating}'s pairwise $F_\mathrm{ST}$ between
all collections, using the {\footnotesize\tt pp.fst} function from the R
package 'hierfstat' \citep{hierfstat}. Finally, we evaluated population
structure in the data by performing unsupervised clustering of
the complete data set (\comment{With or without the LFAR? Which did Anthony use? I would
like to do it with the LFAR.}) 
using the program STRUCTURE \citep{pritchard2000inference,falush2003inference}.
We performed 20 replicate runs at each value of the assumed number of
subpopulations, $K$, in $\{2, 3, 4, 5, 6\}$ 
\comment{I think we should also see what $K=7$ turns up}
using default settings of the program.  STRUCTURE output was
a summarized and visualized using CLUMPAK version~\comment{??}
\citep{kopelman2015clumpak}.

@@

Finally, we investigated the matrix of assignments and misassignments
broken down according to the genotype of the fish at the markers within
the RoSA.  For this, \comment{Need to describe how I categorized
genotypes}.

@@

We tested \comment{XX} candidates from our additional microhaplotype discovery and
retained \comment{106} of those in the panel.

@@

Of these \comment{185} loci that originated from SNPtype assays or from our additional
microhaplotype discovery, \comment{179} of them mapped identically to a single
location on an assembled chromosome in the Otsh\_v2.0 genome, using both bwa mem and BLAT.
Of the remaining 6 loci, one of them was mapped to the same location by BLAT and bwa mem, but
the alignments differed by one base pair in the length of an embedded insertion, two of them had a single secondary alignment on a different chromosome
that was identical in both bwa mem and BLAT, two of them had secondary alignments that
disagreed slightly between BLAT and bwa mem, and one of them had multiple
non-primary alignments (Table~S1 \comment{This is the Excel File.  It needs to
be edited so as to: 1) have a column
giving the category or provenance of the markers (i.e. LFAR or MHAP or SNPTYPE, etc.) 2) retain
only the 2 LFAR markers we actually use. 3) consider renaming the E2 markers since
winter-run-greb1L-diagnostic is confusing...actually, I would consider removing those
entirely---we can introduce them in the subRoSA paper. We don't use them in this paper and I don't
really want to do the explanation of them again.}).

@@

The results are summarized in Fig.~\ref{fig:fprs}.
%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\newcommand{\fprcap}{\footnotesize False positive rates (FPRs) as a function of false negative
rates (FNRs) for distinguishing between different relationships.  {\bf (a)} FPRs calculated
using importance sampling of unrelated pairs of individuals being incorrectly categorized as parent-offspring (PO),
full-sibling (FS) or half-sibling (HS) pairs.  {\bf (b)}  FPRs calculated using regular Monte Carlo
sampling of avuncular pairs (AN: aunt-niece, uncle-nephew, etc.) or full-siblings (FS) or half-siblings (HS) being 
incorrectly categorized as parent-offspring (PO); as well as FPRs of parent-offspring (PO) pairs or half-sibling (HS) 
pairs being incorrectly inferred as full siblings (FS).  In both {\em a} and {\em b}, 100,000 Monte Carlo replicates 
were simulated.  Importance sampling allows estimation of very small false positive rates for unrelated pairs, but 
cannot be applied to truly related pairs.  Consequently, it is difficult to accurately estimate small probabilities 
($10^{-4}$ or less) in {\bf b}, as indicated by the error bars, which represent the estimate $\pm 2s$ where $s$ is the estimated standard error of the mean of the Monte Carlo sample.  This also is the reason that many of the lines terminate.   }
\begin{center}
\includegraphics[width=\textwidth]{images/fpr-fnr-figure-crop.pdf}
\end{center}
\caption[\fprcap]{\fprcap}
\label{fig:fprs}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%
And we can also point readers to the supplement figure \ref{fig:ckmr-comp}.

>>>>

\subsubsection*{Markers potentially associated with phenotypes \citet{thompson2020complex}}

Additionally, we include here genetic markers that characterize variation in two genes, VGLL3 and SIX6,
that have been found to be associated with phenotypes related to reproductive maturity in both
Atlantic \citep{barson2015sex} and Pacific \citep{waters2021heterogeneous} salmonids. We include
them both to provide additional variation for population genetic analyses and to
monitor populations for potential phenotypic patterns.

We include in the reference baseline amplicons targeting SNPs within the ``Region of Strongest Assocation'' (RoSA)
between spring-run and fall-run fish on Chromosome Ots28 described in \citet{thompson2020complex}.  These are useful markers to
have in the panel, as typing them routinely will create a data base for understanding the spatial and temporal
distribution of the alleles at this region; however we have omitted these markers for calculating likelihoods for
GSI, since their distribution across populations in the Feather River has been clearly and evidently
influenced by hatchery practices (see Results), and it is not necessary to include them for accurate assignment.
\comment{Describe additional E1/E2 markers (additional markers within the RoSA region) and
Tasha Marker (because it is close to the duplication noted
by Thompson et al.). We don't go into detail here---keeping it vague because it is
going to be described in the subRoSA paper, but don't mention that.}

We also include a marker that targets the sex-determining region on the Y chromosome in Chinook salmon
and that characterizes genetic sex in Central Valley populations \comment{\citep{von2015development}}

@@

We calculated \citet{weir1984estimating}'s pairwise $F_\mathrm{ST}$ between
all collections, using the {\footnotesize\tt pp.fst} function from the R
package 'hierfstat' \citep{hierfstat}. Finally, we evaluated population
structure in the data by performing unsupervised clustering of
the complete data set
using the program STRUCTURE \citep{pritchard2000inference,falush2003inference}.
We performed 20 replicate runs at each value of the assumed number of
subpopulations, $K$, in $\{2, 3, 4, 5, 6, 7\}$ 
using default settings of the program.  STRUCTURE output was
a summarized and visualized using CLUMPAK version~\comment{??}
\citep{kopelman2015clumpak}.

@@

Finally, we investigated the matrix of assignments and misassignments
broken down according to the genotype of the fish at the markers within
the RoSA.  For this, \comment{Eric. Describe this from what you have in the notebook}.

@@

We tested 229 candidates from our additional microhaplotype discovery and
initially retained the 125 markers described in \citep{thompson2020complex}
for use in stock identification within
the Klamath Basin.  However, further evaluation of those markers in other
populations led us to drop some, and retain 106 of those markers in the current panel.

@@

Of these \comment{185} loci that originated from SNPtype assays or from our additional
microhaplotype discovery, \comment{179} of them mapped identically to a single
location on an assembled chromosome in the Otsh\_v2.0 genome, using both bwa mem and BLAT.
Of the remaining 6 loci, one of them was mapped to the same location by BLAT and bwa mem, but
the alignments differed by one base pair in the length of an embedded insertion, two of them had a single secondary alignment on a different chromosome
that was identical in both bwa mem and BLAT, two of them had secondary alignments that
disagreed slightly between BLAT and bwa mem, and one of them had multiple
non-primary alignments (Table~S1 \comment{This is the Excel File.  It needs to
be edited so as to: 1) have a column
giving the category or provenance of the markers (i.e. LFAR or MHAP or SNPTYPE, etc.) 2) retain
only the 2 LFAR markers we actually use. 3) Definitely rename the E2 markers to be something other
than winter-run-greb1L-diagnostic}).

@@

The results are summarized in Fig.~\ref{fig:fprs}.
%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\newcommand{\fprcap}{\footnotesize False positive rates (FPRs) as a function of false negative
rates (FNRs) for distinguishing between different relationships.  {\bf (a)} FPRs calculated
using importance sampling of unrelated pairs of individuals being incorrectly categorized as parent-offspring (PO),
full-sibling (FS) or half-sibling (HS) pairs.  {\bf (b)}  FPRs calculated using regular Monte Carlo
sampling of avuncular pairs (AN: aunt-niece, uncle-nephew, etc.) or full-siblings (FS) or half-siblings (HS) being 
incorrectly categorized as parent-offspring (PO); as well as FPRs of parent-offspring (PO) pairs or half-sibling (HS) 
pairs being incorrectly inferred as full siblings (FS).  In both {\em a} and {\em b}, 100,000 Monte Carlo replicates 
were simulated.  Importance sampling allows estimation of very small false positive rates for unrelated pairs, but 
cannot be applied to truly related pairs.  Consequently, it is difficult to accurately estimate small probabilities 
($10^{-4}$ or less) in {\bf b}, as indicated by the error bars, which represent the estimate $\pm 2s$ where $s$
is the estimated standard error of the mean of the Monte Carlo sample.  This also is the reason that many of the lines terminate.   }
\begin{center}
\includegraphics[width=\textwidth]{images/fpr-fnr-figure-crop.pdf}
\end{center}
\caption[\fprcap]{\fprcap}
\label{fig:fprs}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%
And we can also point readers to the supplement figure \ref{fig:ckmr-comp}.
The power analyses for this set of markers demonstrate that they have ample variation
for accurate identification of parent-offspring (PO) and full-sibling (FS) pairs in almost all
realistic situations. The predicted false positive rates for both PO and FS pairs are
exceedingly low for even stringent false negative thresholds (e.g. 0.05). 
For example, the false positive rate for FS identification is less than one error in 10 million
comparisons of potential siblings in the California Chinook salmon population with the lowest
heterozygosity (SRW) at a false negative rate of 0.05. However, HS pairs can not be accurately
distinguished from unrelated pairs in any populations without very large false negative rates
and in very modest-sized studies.
We note, however, that the presence in a dataset of genotypes from individuals that are related, but not in
 the relationship category proposed, will result in some additional misidentifications.
 These 'errors' will all involve related fish being identified as such, but in the incorrect relationship category.  The cross-assignment frequencies predicted for some common relationship categories, and for realistic FPR and FNR, are found in (Fig 4b).

\textbf{Although some additional errors might be
expected when related but non-FS individuals are sampled,  this number of comparisons exceeds
those necessary in almost all realistic scenarios.
We note that, as in previous work, it is nearly impossible to identify HS pairs accurately
with data from this, or any standard marker dataset. However, the use of properly LD-filtered
genome-scale data should allow such discrimination in future work.
Accurate relationship
inference in even the most genetically depauperate Chinook salmon population in California
and rangewide (Seeb et al. 2007, Clemento et al. 2014) is therefore possible with this marker set.}

While there is not a currently available method (like that of \citeauthor{anderson2006power}) for estimating error rates in
inference of parent-offspring trios with multiallelic marker data, given the substantial amount of statistical power for identifying PO and FS pairs, which require much more statistical power, it is clear that PO-trio inference with these markers will be essentially error free.
